---
title: "String Manipulation with stringr"
author: "Aditya Dube"
date: "`r format(Sys.time(), '%A, %B %e, %Y')`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Loading Packages

```{r }
library(tidyverse)
library(skimr)
library(stringr)
library(lubridate)
```


# Character Strings

## Importing Dataset 

```{r }
# Variables to keep
keeps <- c("track_name", "youtube_title", "youtube_duration", "full_lyrics")

# Importing CSV file
swiftSongs <- read_csv("https://raw.githubusercontent.com/dilernia/STA418-518/main/Data/swiftSongs.csv") %>% select(keeps)

```


##  Explore high-level characteristics of the data using the glimpse() and skim() functions.

```{r }
glimpse(swiftSongs)
skim(swiftSongs)
```


# Matching Strings

```{r }
# Displaying lyrics
swiftSongs %>% filter(track_name == "It’s Nice To Have A Friend") %>% 
  pull(full_lyrics)

```

```{r }
# Detecting if a string contains the substring 'Taylor'
str_detect(string = c("Taylor Swift", "Taylor Lautner", "Harry Styles"),
           pattern = "Taylor")

```


## Using the str_detect() and mutate() functions, add a new boolean variable called contains_midnight to swiftSongs that indicates whether or not a song’s lyrics contain the word “midnight”. 

```{r }
swiftSongs <- swiftSongs %>% mutate(contains_midnight = str_detect(full_lyrics, "\\bmidnight\\b")) 

```

## How many of Taylor’s songs mention the word “midnight”?
## Answer: 5

```{r }
sum(swiftSongs$contains_midnight)
```
## How many of Taylor’s songs mention the word “midnight” or “Midnight”?
## Answer : 6

```{r }
swiftSongs <- swiftSongs %>%
  mutate(contains_midnight_or_Midnight = str_detect(full_lyrics, regex("\\bmidnight\\b|\\bMidnight\\b")))

sum(swiftSongs$contains_midnight_or_Midnight)
```
## Using the str_count() and mutate() functions, add a new variable called love_count to swiftSongs that indicates how many times each song mentions the word “love”.

```{r }

swiftSongs <- swiftSongs %>% mutate(love_count = str_count(full_lyrics, pattern = "love"))
```


## Which song mentions love the most times, and how many times is it mentioned?

```{r }
song_with_max_love <- swiftSongs %>%
  filter(love_count == max(love_count)) %>%
  slice_max(love_count)

# Extract the song name and the love count
song_name <- song_with_max_love$youtube_title
max_love_count <- song_with_max_love$love_count

song_name
max_love_count

```


# Modifying Strings

## Create a new variable called youtube_time that is the same as youtube_duration, but with a : symbol replacing the M.

```{r }

swiftSongs <- swiftSongs %>%
  mutate(youtube_time = str_replace(youtube_duration, "M", ":"))


```


## Modify youtube_time by removing the P, T, and S letters.

```{r }
swiftSongs <- swiftSongs %>%
  mutate(youtube_time = str_remove_all(youtube_time, pattern = "P|T|S"))

```


##  Modify youtube_time to add  0’s when needed using the case_when() function together with str_replace_all() and str_length().

```{r }
swiftSongs <- swiftSongs %>%
  mutate(youtube_time = case_when(
    str_length(youtube_time) == 2 ~ str_c(youtube_time, "00"),
    str_length(youtube_time) == 3 ~ str_replace_all(youtube_time, pattern = ":", replacement = ":0"),
    TRUE ~ youtube_time
  ))
```


## Coerce youtube_time to be a special date / time variable using the parse_date_time() function from the lubridate package using the code below

```{r }
# Coercing youtube_time to a date / time variable
swiftSongs <- swiftSongs %>% 
  dplyr::mutate(youtube_time = lubridate::parse_date_time(youtube_time, orders = "%M:%S"))

```


## Use the minute() and second() functions from the lubridate package, create a new variable, song_duration_s that gives the song duration in seconds using the code below

```{r }
# Creating song_duration_s variable
swiftSongs <- swiftSongs %>% 
  dplyr::mutate(song_duration_s = lubridate::second(youtube_time) + 
                               60*lubridate::minute(youtube_time))
```

## The escape sequence \\w+ can be used to match any ‘word’ character (although it very slightly over counts). Create a new variable song_words equal to the number of words in the song using the str_count() function and the full_lyrics variable.

```{r }
# Creating song_words variable
swiftSongs <- swiftSongs %>% 
  dplyr::mutate(song_words = str_count(full_lyrics, pattern = "\\w+"))

```


## Reproduce the plot below showing the relationship between the duration of each song in seconds and its number of words. Hint: to match the style of the points, use fill = '#01a7d9', pch = 23, color = '#7d488e' inside of the geom_point() layer.

```{r }
ggplot(data = swiftSongs, aes(x= song_duration_s,y= song_words)) + geom_point(fill = '#01a7d9', pch = 23, color = '#7d488e') + theme_bw() + labs(
  title ="Number of words by Taylor Swift song duration",
  x = "Song duration (seconds)",
  y = "Number of words in lyrics",
  caption = "Data source: geniusr R package"
)

```



# Capitalization and spacing

```{r }
# Setting all characters to lowercase
str_to_lower("It’s nice to have a friend")
## [1] "it’s nice to have a friend"
# Setting all characters to uppercase
str_to_upper("It’s nice to have a friend")
## [1] "IT’S NICE TO HAVE A FRIEND"
# Setting all characters to title case
str_to_title("It’s nice to have a friend")

```


```{r }
# Removing spaces at start and end of string
str_trim(" Best believe I'm still bejeweled     When I walk in the room     I can still make the whole place shimmer ")
## [1] "Best believe I'm still bejeweled     When I walk in the room     I can still make the whole place shimmer"
# Removing spaces at start and end of string and repetitive spaces
str_squish(" Best believe I'm still bejeweled     When I walk in the room     I can still make the whole place shimmer ")

```


# Bonus (optional)

## Using the title of the YouTube video for each song, create a variable indicating whether or not the video is an official music video, official lyric video, or other type of video.


```{r }
swiftSongs <- swiftSongs %>%
  mutate(video_type = case_when(
    str_detect(youtube_title, "(?i)official\\s+music\\s+video") ~ "Official Music Video",
    str_detect(youtube_title, "(?i)official\\s+lyric\\s+video") ~ "Official Lyric Video",
    TRUE ~ "Other"
  ))

```


## Use the str_glue() function in tandem with ggplot to create a scatter plot showing the relationship between the total number of characters in each song’s lyrics (full_lyrics) and the total number of characters in each song’s title (track_name), including the correlation between the two variables dynamically in the subtitle.

```{r }
swiftSongs <- swiftSongs %>%
  mutate(total_lyrics_chars = str_length(full_lyrics),
         total_title_chars = str_length(youtube_title))

# Calculate the correlation between the two variables
correlation <- cor(swiftSongs$total_lyrics_chars, swiftSongs$total_title_chars)

# Create the scatter plot with dynamic correlation subtitle
ggplot(swiftSongs, aes(x = total_lyrics_chars, y = total_title_chars)) +
  geom_point() +
  labs(x = "Total Lyrics Characters", y = "Total Title Characters",
       subtitle = str_glue("Correlation: {correlation}")) +
  theme_bw()


```

# Making a ☁️Word Cloud☁️

```{r }
library(wordcloud2) 
library(tidytext)

# Tallying up frequency of words in all songs
wordFreqs <- swiftSongs %>% unnest_tokens(word, full_lyrics) %>% 
  count(word) 

# Removing 'stop words' (common but not very meaningful words)
wordFreqs <- wordFreqs %>% anti_join(stop_words)

# Creating word cloud
wordcloud2(wordFreqs, size=1.6, color='random-dark',shape = 'star')

```

