---
title: "Homework 4: The NHANES Data Set"
author: "Aditya Dube"
date: "`r format(Sys.time(), '%A, %B %e, %Y')`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Question 1

## Part A

```{r }
library(tidyverse)
library(skimr)
```

## Part B

```{r }
# Loading the NHANES data set
data(NHANES, package = "NHANES")
```


# Question 2

## Part A

```{r }
glimpse(NHANES)
```


## Part B

## There are 10,000 rows and 76 columns in the dataset.


## Part C

## Based on the output obtained from glimpse(), which type / class of variable is Age? What about BMI and Diabetes?

## Answer: Age is of int type and BMI is dbl and Diabetes is fct. 


# Question 3


## Part A

```{r }
skim(NHANES)
```



## Part B

## Answer: In fct variable - 'BMICatUnder20yrs' has the most missing values and in numeric variable - 'HeadCirc' has the most missing values

## Part C

```{r }
NHANES_sub <- select(NHANES, 1:5)
glimpse(NHANES_sub)
skim(NHANES_sub)
```

## Part D

```{r }
count(NHANES, SmokeNow, Diabetes)
```

## Part E

```{r }
#Total diabetic people are 760
sum(NHANES$Diabetes == "Yes", na.rm = TRUE) 

#Total people are not diabetic 9098
sum(NHANES$Diabetes == "No", na.rm = TRUE)
```

## Part F

```{r }
#Missing values are there for diabetes status - 142
sum(is.na(NHANES$Diabetes))
```

## Part G

```{r }
#Percentage of people with diabetes among smokers
sum(NHANES$SmokeNow == "Yes" & NHANES$Diabetes == "Yes", na.rm = TRUE) /
                           sum(NHANES$SmokeNow == "Yes" & !is.na(NHANES$Diabetes), na.rm = TRUE) * 100


#Percentage of people with diabetes among non-smokers
sum(NHANES$SmokeNow == "No" & NHANES$Diabetes == "Yes", na.rm = TRUE) /
                               sum(NHANES$SmokeNow == "No" & !is.na(NHANES$Diabetes), na.rm = TRUE) * 100

```

## Part H

## Answer: There is a lower prevalence of diabetes among smokers compared to non-smokers in the given dataset. This result contradicts the general understanding that smoking is associated with an increased risk of diabetes. In most studies, smoking is indeed considered a risk factor for developing diabetes.



# Question 4

```{r }
# Printing rows with 5 lowest BMIs
slice_head(arrange(NHANES, BMI), n = 5)
```

```{r }
# Printing rows with 5 lowest BMIs
NHANES %>% arrange(BMI) %>% slice_head(n = 5)
```


## Part A


```{r }
NHANES %>% arrange(BMI) %>% slice_tail(n = 5)
```

## Part B

```{r}
NHANES %>% slice_min(BMI, n = 5) # This one is printing more than 5 rows because of the duplicate value or ties

NHANES %>% slice_min(BMI, n = 5, with_ties = FALSE) # Now this is printing 5 lowest BMI
NHANES %>% slice_max(BMI, n = 5) # 5 highest BMI
```

## Part C

## Answer: People with high BMI are overweight,  Higher BMI values is more common in older age groups and also BMI values may vary based on socioeconomic factors, such as access to healthy food, education level, or cultural influences and Lower BMI values are relatively healthier in terms of weight-related conditions.