---
title: "Intermediate ggplot2"
author: "Aditya Dube"
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
date: "`r format(Sys.time(), '%A, %B %e, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Loading Packages

```{r }
library(riskCommunicator)
library(tidyverse)
library(skimr)
library(knitr)
library(ggthemes)
library(patchwork)
```

## First, let’s load the FHS data set from the riskCommunicator package

```{r }
data(framingham, package = "riskCommunicator")
glimpse(framingham)

```

## Select the first 10 variables from the Framingham dataset and store it as a new data frame called framinghamSub using the select() function. Also, update the SEX variable to have the values “Male” and “Female” rather than 1 and 2, and the CURSMOKE variable to have the values “Yes” and “No” rather than 1 and 0 using the mutate() and case_when() functions. This should be your new dataset to be used for the rest of the assignment.


```{r }
framinghamSub <- framingham %>% select(1:10) %>%  mutate(
  SEX = case_when(SEX == 1 ~ "Male",
                  SEX == 2 ~ "Female",
                  TRUE ~ as.character(SEX)),
  CURSMOKE = case_when(
    CURSMOKE == 1 ~ "Yes",
    CURSMOKE == 0 ~ "No",
    TRUE ~ as.character(CURSMOKE)
  )
)
```

## Use the skim() function from the skimr package to explore other characteristics of the subset of the data.

```{r }
skim(framinghamSub)
```


## Make a scatter plot between diastolic (DIABP) and systolic (SYSBP) blood pressure with a “facet” by the sex of the participant (SEX). Also manually set the alpha aesthetic to be 0.2. After the next few bullets is an example

```{r }
ggplot(data = framinghamSub, mapping = aes(x = SYSBP, y = DIABP)) + geom_point(alpha = 0.20) + facet_grid(. ~ SEX)

```

## Also include the size of the data points as mapped by the number of cigarettes smoked per day (CIGPDAY), add a color-blind friendly palette for coloring the points based on the sex of each participant, and position the legend at the bottom of the plot.

```{r }
ggplot(data = framinghamSub,
       mapping = aes(
         x = SYSBP, y = DIABP,
         size = CIGPDAY,
         color = SEX
       )) + geom_point(alpha = 0.20) + facet_grid(. ~ SEX) + scale_color_viridis_d() + labs(
         title = "Systolic by diastolic blood pressure" ,
         x = "Systolic BP" ,
         y = "Diastolic BP",
         caption = "Data source: Framingham Heart Study & the riskCommunicator package "
       ) + theme(legend.position = "bottom")
```


## Add a line of best fit corresponding to a simple linear regression model fit separately for males and females using geom_smooth().


```{r }
scatter <- ggplot(data = framinghamSub,
       mapping = aes(
         x = SYSBP,
         y = DIABP,
         size = CIGPDAY,
         color = SEX
       )) + geom_point(alpha = 0.20) + facet_grid(. ~ SEX) + scale_color_viridis_d() + labs(
         title = "Systolic by diastolic blood pressure" ,
         x = "Systolic BP" ,
         y = "Diastolic BP",
         caption = "Data source: Framingham Heart Study & the riskCommunicator package ", size = "Cigarettes smoked per day"
       ) + guides(color = FALSE) + geom_smooth(se = FALSE, method = "lm", size = 1) + theme_bw() + theme(legend.position = "bottom")

scatter
```

## Next, create a side-by-side box-plot where the y-axis is total cholesterol (TOTCHOL) and the x-axis is current smoking status (CURSMOKE). Make all axis and title text bold in the plot.

```{r }
ggplot(
  data = framinghamSub,
  mapping = aes(
    x = CURSMOKE,
    y = TOTCHOL,
  )
) + geom_boxplot() + labs(
  title = "Total cholesterol by smoking status" ,
  x = "Current smoker" ,
  y = "Serum Total Cholesterol (mg/dL)",
  caption = "Data source: Framingham Heart Study & the riskCommunicator package "
)  +
  theme(
    text = element_text(face = "bold"))

```



## Add a complete theme from ggthemes, color the boxes based on smoking status, remove the legend, and make the axis titles bold and change the font size as well.

```{r }
ggplot(
  data = framinghamSub,
  mapping = aes(
    x = CURSMOKE,
    y = TOTCHOL,
    fill = CURSMOKE
  )
) + geom_boxplot() + scale_fill_viridis_d() + labs(
  title = "Total cholesterol by smoking status" ,
  x = "Current smoker" ,
  y = "Serum Total Cholesterol (mg/dL)",
  caption = "Data source: Framingham Heart Study & the riskCommunicator package "
)   + theme_economist() +
  theme(
    text = element_text(face = "bold"), legend.position = "none",  axis.title = element_text(size = 14))

```


## In a new plot, modify the side-by-side box-plots we created to be faceted by the sex of the participant using the facet_grid() function and columns to break up the subplots.

```{r }
ggplot(
  data = framinghamSub,
  mapping = aes(
    x = CURSMOKE,
    y = TOTCHOL,
    fill = CURSMOKE
  )
) + geom_boxplot() + facet_grid(. ~ SEX) + scale_fill_viridis_d() + labs(
  title = "Total cholesterol by smoking status" ,
  x = "Current smoker" ,
  y = "Serum Total Cholesterol (mg/dL)",
  caption = "Data source: Framingham Heart Study & the riskCommunicator package "
)   + theme_economist() +
  theme(
    text = element_text(face = "bold"), legend.position = "none",  axis.title = element_text(size = 14))
```


## Make a line graph that shows the average cigarettes per day (CIGPDAY) by age (AGE), with separate lines by the sex of the participant (SEX).


```{r }
framinghamSub %>% ggplot() + stat_summary(aes(x= AGE, y= CIGPDAY, group = SEX, color = SEX), geom = "line", size = 1, fun.y = mean)
```


## Apply a complete theme to the plot, and have the axis show the breaks at 0, 4, 8, 12, and 16 cigarettes per day.

```{r }
lineChart <- framinghamSub %>% ggplot() + stat_summary(aes(x= AGE, y= CIGPDAY, group = SEX, color = SEX), geom = "line", size = 1, fun.y = mean) +
  labs(title = "Average cigarettes per day by age and sex" ,
  x = "Age" ,
  y = "Average cigarettes per day",
  color = "Sex") + scale_color_colorblind() + scale_y_continuous(breaks=c(0,4,8,12,16)) + theme_bw()

lineChart
```


## Combine the line chart and the faceted scatter plots together into a single graphic using the patchwork package, with 1 plot per row and the line chart on top.


```{r}
lineChart / scatter
```



## Bin / categorize total cholesterol levels is as Normal (<200 mg/dL), Borderline high (200 to 239 mg/dL), or High (> 240 mg/dL). 

```{r }
framinghamSub <- framinghamSub %>% 
  mutate(CholesterolCat = case_when(TOTCHOL < 200 ~ "Normal",
                                    TOTCHOL >= 200 &  TOTCHOL < 240 ~ "Borderline high",
                                    TOTCHOL > 240 ~ "High",
                         TRUE ~ as.character(NA)))

```


## Create a bar chart displaying the number of participants falling in each cholesterol category based on Johns Hopkins’ definitions using geom_bar(). Also, remove people under 40 and those without recorded cholesterol levels (missing values for CholesterolCat) from the plot by using the code filter(AGE >= 40, !is.na(CholesterolCat)) when piping the data into each subsequent ggplot() call.


```{r }
framinghamSub %>%  filter(AGE >= 40, !is.na(CholesterolCat)) %>% ggplot(aes(x=CholesterolCat)) + geom_bar()

```

## Recreate the bar chart, this time reordering the categories to show Normal, Borderline high, and then High from left to right using the fct_relevel() function.

```{r }
framinghamSub %>%  filter(AGE >= 40, !is.na(CholesterolCat)) %>% ggplot(aes(x=fct_relevel(CholesterolCat, "Normal", "Borderline high", "High"))) + geom_bar() 

```


## Change the color of the inside of the bars based on the cholesterol category using a color-blind friendly palette, make the outline of the bars black in color, facet by the sex of the participant, and remove the legend.


```{r }
framinghamSub %>%  filter(AGE >= 40, !is.na(CholesterolCat)) %>% ggplot(aes(x=fct_relevel(CholesterolCat, "Normal", "Borderline high", "High"), fill = CholesterolCat ))  +
  labs(title = "Distribution of cholesterol levels, Age 40+" ,
  x = "Cholesterol level" ,
  y = "Count", 
  caption = "Data source: Framingham Heart Study & the riskCommunicator package") + facet_grid(. ~ SEX) + scale_fill_viridis_d()+ theme_bw()  + geom_bar(color = "black") + theme(legend.position = "none")

```

## Lastly, use the coord_flip() function to turn the bar chart into a horizontal bar chart instead

```{r }
framinghamSub %>%  filter(AGE >= 40, !is.na(CholesterolCat)) %>% ggplot(aes(x=fct_relevel(CholesterolCat, "Normal", "Borderline high", "High"), fill = CholesterolCat ))  +
  labs(title = "Distribution of cholesterol levels, Age 40+" ,
  x = "Cholesterol level" ,
  y = "Count", 
  caption = "Data source: Framingham Heart Study & the riskCommunicator package") + facet_grid(. ~ SEX) + scale_fill_viridis_d()+ theme_bw()  + geom_bar(color = "black") + theme(legend.position = "none") + coord_flip()
```
