---
title: "Fundamentals of Functions in R"
author: "Aditya Dube"
date: "`r format(Sys.time(), '%A, %B %e, %Y')`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Loading Packages

```{r }
library(tidyverse)
library(stringr)
```

# Example function

## Create the mySum() function, and use it to calculate the sum of 13 and 1989.

```{r }
mySum <- function(number1, number2) {
  return(number1 + number2)
}

# Defaulting to order of arguments in function
mySum(13, 1989)
```
## What happens when we calculate the sum of NA and 1989?
## Answer: It is NA

```{r }
mySum(NA, 1989)
```
# Best practices

## Copy the code creating the myNewSum() function below and add comments to communicate the purpose of different parts of the code.

```{r }
# The function myNewSum calculates the sum of two numbers, number1 and number2.
# It has an optional argument remove.na which, if set to TRUE, replaces any NA values with 0 before performing the sum.

myNewSum <- function(number1, number2, remove.na = TRUE) {
  # Check if remove.na is set to TRUE
  if(remove.na) {
    # If remove.na is TRUE, replace NA values with 0 using ifelse() function
    # Add number1 and number2 after replacing NA values with 0
    calcSum <- ifelse(is.na(number1), 0, number1) + 
               ifelse(is.na(number2), 0, number2)
  } else {
    # If remove.na is FALSE, simply add number1 and number2
    calcSum <- number1 + number2
  }
  
  # Return the calculated sum
  return(calcSum)
}

myNewSum(1,3)
```

##  Read the source code for each of the following three functions, determine the purpose of each function, and propose better function names.

## Answer: The first i will be changing is the naming convention f1, f2 and f3 are poor naming convention

```{r }
startsWithPrefix <- function(string, prefix) {
  str_sub(string, 1, nchar(prefix)) == prefix
}
startsWithPrefix("Hello, world!", "Hello") 

removeLastElement <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
removeLastElement(c(1, 2, 3, 4))

harmonicMean <- function(x) {
  1 / (mean(1 / x))
}
harmonicMean(c(2, 4, 6)) 
```

## Create a new function to calculate the geometric mean of a vector of numbers which is equal to the nth root of the product of n numbers.

```{r }
geometricMean <- function(x) {
  if (length(x) == 0) {
    stop("Input vector must not be empty.")
  }
  
  if (any(x <= 0)) {
    stop("Input vector must contain positive numbers only.")
  }
  
  n <- length(x)
  geom_mean <- prod(x)^(1/n)
  
  return(geom_mean)
}

geometricMean(c(5,2))
```

# Conditional execution, warnings, and errors

## Extend / modify the calcSqrt() function to have an error instead of a warning occur if the input number is negative. Use calcSqrt() to try to calculate the square root of 4 and -9.

```{r error=TRUE}
calcSqrt <- function(val) {
  if (val >= 0) {
    return(val^0.5)
  } else {
    stop("val is negative but must be >= 0!")
    return(NaN)
  }
}
calcSqrt(4)
calcSqrt(-9)
```

##  Create a function called is.whole.number(), which checks if a given value is a whole number (e.g., 0, 1, 2, …) or not.

```{r }
is.whole.number <- function(value) {
  if (is.integer(value) || (is.numeric(value) && value %% 1 == 0)) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}
is.whole.number(4)    # Returns TRUE
is.whole.number(3.0)  # Returns TRUE
is.whole.number(3.5)  # Returns FALSE
```
## The ^ operator raises a number to a specified power in R. Create a function that raises a number to a specified power, and returns a warning and an NaN value if the number is negative and the power is not a whole number.

```{r }
raiseToPower <- function(base, exponent) {
  if (base >= 0 || is.whole.number(exponent)) {
    return(base^exponent)
  } else {
    warning("Base is negative and exponent is not a whole number!")
    return(NaN)
  }
}
# Use your function to calculate 4^(2), (-9)^(2), 3^(1/3), and (-3)^(1/3).
raiseToPower(4, 2)
raiseToPower(-9, 2)
raiseToPower(3, (1/3))
raiseToPower(-3, (1/3))
```
# The ifelse() function

## Create the myNewSum() function, and use it to calculate the sum of 13 and 1989

```{r }
# Creating a function called mySum with 2 required arguments, 1 optional argument
myNewSum <- function(number1, number2, remove.na = TRUE) {
  if(remove.na) {
    calcSum <- ifelse(is.na(number1), 0, number1) + 
         ifelse(is.na(number2), 0, number2)
  } else {
    calcSum <- number1 + number2
  }
  return(calcSum)
}
```

## What happens when we calculate the sum of NA and 1989 using myNewSum()?

```{r }
myNewSum(NA, 1989)
#It return 1989 
```
# Logical operators

##  Create a function that takes a numeric vector as input, and returns a single boolean value indicating whether or not the vector has any negative values.

```{r }
hasNegativeValues <- function(vector) {
  any(vector < 0)
}
hasNegativeValues(c(1, 2, -3, 4))
```

## Does the code (1:4 < 4) & (1:4 > 1) produce the same result as (1:4 < 4) && (1:4 > 1)? Which code achieves what appears to be its intended purpose?

## Answer: No, the code (1:4 < 4) & (1:4 > 1) does not produce the same result as (1:4 < 4) && (1:4 > 1). The & operator performs element-wise logical AND operation on the corresponding elements of the two vectors, while the && operator performs a single logical AND operation considering only the first element of each vector.


# Bonus (optional)

## Write a function to check if a number is even or odd that returns a vector containing the values “even”, “odd”, or “not an integer” depending on its input.

```{r }
checkEvenOdd <- function(nums) {
  result_x <- sapply(nums, function(num) {
    if (is.whole.number(num)) {
      if (num %% 2 == 0) {
        return("even")
      } else {
        return("odd")
      }
    } else {
      return("not an integer")
    }
  })
  return(result_x)
}

checkEvenOdd(c(5, 6, 7, 8))


```
