---
title: "Homework 2"
author: "Aditya Dube"
date: "`r format(Sys.time(), '%A, %B %e, %Y')`"
output: html_document
---

# Question 1 Parts 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

## Loading Packages 

```{r message = FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(knitr)
library(arrow)
library(bench)
library(ggbeeswarm)
library(haven)
```
# Question 2 Parts 

## Importing Dataset 

```{r }
swiftSongs = read.csv(
  "https://raw.githubusercontent.com/dilernia/STA418-518/main/Data/swiftSongs.csv"
)
```

## Print a single table of the swiftSongs data containing the last 13 rows of the data using the slice_tail() function and the 30th through 34th columns using the select() function from the dplyr package.

```{r }
swiftSongs %>% dplyr::select(30:34) %>% 
  slice_tail(n = 13) %>% kable()
```

## Import data on Taylor Swift songs directly from the URL using the read.csv(), read_csv(), and the fread() functions, comparing the read times using the mark function from the bench package, storing the results of the mark function in an object called readTimes. Specify a minimum of 5 iterations in the mark function.

```{r }
readTimes <-
  mark(
    read.csv(
      "https://raw.githubusercontent.com/dilernia/STA418-518/main/Data/swiftSongs.csv"
    ),
    read_csv(
      "https://raw.githubusercontent.com/dilernia/STA418-518/main/Data/swiftSongs.csv",
      show_col_types = FALSE
    ),
    fread(
      "https://raw.githubusercontent.com/dilernia/STA418-518/main/Data/swiftSongs.csv"
    ),
    check = FALSE,
    min_iterations = 5
  )
ggObj <- plot(readTimes)

importTimes <- ggObj$data %>% mutate(expression =
                                       paste0(map_chr(str_split(expression, pattern = "[(]"), 1), "()"))
# Printing table
importTimes %>% arrange(desc(median)) %>%
  select(expression:mem_alloc) %>% distinct() %>% knitr::kable()

```

## Create a violin plot to display the varying speeds of the three different functions for importing the data. 

```{r }
importTimes %>% ggplot(aes(x = time, y = fct_reorder(expression, time))) +
  geom_violin(fill = "dodgerblue") +
  geom_jitter(
    height = 0.2,
    pch = 21,
    fill = "black",
    color = "white"
  ) +
  labs(title = "Comparison of read times", y = "", x = "Run time") +
  theme_bw()
```

## Based on the violin plot, which function was typically the fastest?

## Answer: fread() function was the fastest.

# Question 3 Parts

## Export data on Taylor Swift songs in the swiftSongs data set using the write.csv(), write_csv(), and the fwrite() functions, comparing the write times using the mark function from the bench package, storing the results of the mark function in an object called writeTimes. Specify a minimum of 5 iterations in the mark function.

```{r }

writeTimes <-
  mark(
    write.csv(swiftSongs, "swiftSongs.csv"),
    write_csv(swiftSongs, "swiftSongs.csv"),
    fwrite(swiftSongs, "swiftSongs.csv"),
    min_iterations = 5,
    check = FALSE,
    memory = FALSE
  )

ggObj <- plot(writeTimes)

exportTimes <- ggObj$data %>% mutate(expression =
                                       paste0(map_chr(str_split(
                                         expression, pattern = "[(]"
                                       ), 1), "()"))
# Printing table
exportTimes %>% arrange(desc(median)) %>%
  select(expression:mem_alloc) %>% distinct() %>% knitr::kable()
```
## Create a violin plot to display the varying speeds of the three different functions for exporting the data. 

```{r }
exportTimes %>% ggplot(aes(x = time, y = fct_reorder(expression, time))) +
  geom_violin(fill = "dodgerblue") +
  geom_jitter(
    height = 0.2,
    pch = 21,
    fill = "black",
    color = "white"
  ) +
  labs(title = "Comparison of write times", y = "", x = "Run time") +
  theme_bw()
```

## Based on the violin plot, which function was typically the fastest?

## Answer: fwrite() is the fastest