---
title: "Final Project Activity"
author: "Aditya Dube"
date: "`r format(Sys.time(), '%A, %B %e, %Y')`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Packages
```{r }
library(tidyverse)
library(skimr)
library(naniar)
library(ggplot2)
library(dplyr)
library(maps)
library(flextable)
library(plotly)
```
# Q1. Import your data into R.

# Importing Fast Food Dataset and Census Dataset

```{r}
fastFoodData <-  read_csv("fastFoodGeo.csv")
```

```{r }
censusData <- read_csv("census_data_full_2008-2021.csv")
```

# Merging both the datset 
```{r }
mergedData <- inner_join(fastFoodData, censusData, by = "geoid")
```

# Q2. Explore and display high-level characteristics of your data set, e.g., important variables and their types, different levels for factor variables, any patterns of missing values.

```{r }
# Display the variable names
names(mergedData)
```

```{r }
#Broad overview of a data frame
skim(mergedData)
```

```{r}
# Display the structure of the dataset
str(mergedData)
```
```{r }
# Display summary statistics of the dataset
summary(mergedData)

```

```{r }
# Check for missing values
missing_values <- sum(is.na(mergedData))

# Creating a bar plot showing the proportion of missing values for each variable
gg_miss_var(mergedData)
```



## Cleaning The Dataset After Merging

# Removing Variables related to Race and Education as there were many missing datas and also removing Website variable.

# Keeping Variables
```{r }
cleanedData <- mergedData %>% select(geoid,keys,address,city,country,name,postalCode,province,lat,long,year,population,median_income,median_monthly_rent_cost,median_monthly_home_cost, prop_female,prop_male, prop_poverty)
```

```{r }
# Display the structure of the dataset
str(cleanedData)
```

```{r }
# Display summary statistics of the dataset
summary(cleanedData)
```

```{r }
# Check for missing values
missing_values_cleaned_data <- sum(is.na(cleanedData))

# Creating a bar plot showing the proportion of missing values for each variable
gg_miss_var(cleanedData)
```

```{r }
#Broad overview of a data frame
skim(cleanedData)
```


# Q3. Obtain at least one table of group-level summary statistics for quantitative variables in your data set


```{r }
# Group-level summary statistics for 'median_income' by 'province'
summary_table <- cleanedData %>%
  group_by(province) %>%
  summarise(mean_median_income = mean(median_income),
            median_median_income = median(median_income),
            sd_median_income = sd(median_income),
            min_median_income = min(median_income),
            max_median_income = max(median_income))

# Keep only the rows with the maximum values of 'mean_median_income' for each province
max_mean_income <- summary_table %>%
  slice_max(n = 1, order_by = mean_median_income)

# Display the row(s) with the maximum mean median income
max_mean_income
```
# Q4. Obtain at least one frequency table showing the cross-section of two categorical variables in your data set. 

```{r }
# Cross-tabulation of 'city' and 'province'
frequency_table <- table(cleanedData$city, cleanedData$province)

# Convert the frequency table to a data frame
frequency_df <- as.data.frame(frequency_table, stringsAsFactors = FALSE)

# Keep only the rows with the highest frequencies for each city
# Var1 is Cities Names
top_frequencies <- frequency_df %>%
  group_by(Var1) %>%
  slice_max(n = 1, order_by = Freq)

# Display the top frequencies
top_frequencies
```

# Q5. Create at least one ggplot visualization for your data, including appropriate labels with units of measurement when applicable and a specified theme

```{r }
# Create a bar plot of population by province
bar_plot <- ggplot(data = cleanedData, aes(x = province, y = population)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Province", y = "Population", title = "Population by Province") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Convert the bar plot to an interactive plot using plotly
interactive_plot <- ggplotly(bar_plot)

# Display the interactive plot
interactive_plot
```

# Q6. Create a data dictionary showcasing the variables used in your analyses. An example R Markdown file showing creation of a data dictionary for the Palmer Penguins data set is available on Blackboard

```{r }
dataDictionary <- tibble(Variable = colnames(cleanedData),
                         Description = c("Geographic Region ID",
                                         "Keys for identification purposes",
                                         "Address of the location",
                                         "City where the location is situated",
                                         "Country of the location",
                                         "Name of the location",
                                         "Postal code of the location",
                                         "Province or state where the location is located",
                                         "Latitude of the location",
                                         "Longitude of the location",
                                         "Year of data collection",
                                         "Population count",
                                         "Median income of the location",
                                         "Median monthly rental cost of housing",
                                         "Median monthly cost of home ownership",
                                         "Proportion of population that is female",
                                         "Proportion of population that is male",
                                         "Proportion of people 25 and older living in poverty"),
                         Type = map_chr(cleanedData, .f = function(x){typeof(x)[1]}),
                         Class = map_chr(cleanedData, .f = function(x){class(x)[1]}))

#Setting theme for flextable
set_flextable_defaults(
  font.size = 10, theme_fun = theme_vanilla,
  padding = 6,
  background.color = "#EFEFEF")

flextable::flextable(dataDictionary, cwidth = 2)
```

