---
title: "Introduction to ggplot2"
author: "Aditya Dube"
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
date: "`r format(Sys.time(), '%A, %B %e, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE, error = TRUE)
```

## Loading Packages

```{r}
library(palmerpenguins)
library(tidyverse)
library(knitr)
library(dplyr)
library(skimr)
```

## Loading Dataset

```{r }
data(penguins, package = "palmerpenguins")
glimpse(penguins)
```

## First Start with blank canvas

```{r }
ggplot(data = penguins)
```

## Next add an x-axis variable

```{r }
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm))

```

## After that adding variable to y-axis.

```{r }
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm, y = bill_length_mm))
```

## Create a Scatter Plot for bill_depth_mm and bill_length_mm

```{r }
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm, y = bill_length_mm)) + geom_point()
```

# Aesthetics for Scatter Plot

## Adding color
```{r }
ggplot(
  data = penguins,
  mapping = aes(x = bill_depth_mm, y = bill_length_mm, color = species)
) + geom_point() 
```

## Adding title and subtitle

```{r }
ggplot(
  data = penguins,
  mapping = aes(x = bill_depth_mm, y = bill_length_mm, color = species)
) + geom_point() + labs(title = "Penguin bill length by depth", subtitle = "Dimensions for Adelie, Chinstrap, Gentoo penguins")
```

## Customize the axis labels using the x and y options

```{r }
ggplot(
  data = penguins,
  mapping = aes(x = bill_depth_mm, y = bill_length_mm, color = species)
) + geom_point() + labs(title = "Penguin bill length by depth", subtitle = "Dimensions for Adelie, Chinstrap, Gentoo penguins", x = "Bill depth (mm)" , y = "Bill length (mm)")
```

## Customizing the legend title and plot caption


```{r }
ggplot(
  data = penguins,
  mapping = aes(x = bill_depth_mm, y = bill_length_mm, color = species)
) + geom_point() + labs(title = "Penguin bill length by depth", subtitle = "Dimensions for Adelie, Chinstrap, Gentoo penguins", x = "Bill depth (mm)" , y = "Bill length (mm)", caption = "Source: Palmer Station LTER/palmerpenguins package", color = "Species")
```

## Lastly, we use a color blind friendly palette.

```{r }
ggplot(
  data = penguins,
  mapping = aes(x = bill_depth_mm, y = bill_length_mm, color = species)
) + geom_point() + labs(title = "Penguin bill length by depth", subtitle = "Dimensions for Adelie, Chinstrap, Gentoo penguins", x = "Bill depth (mm)" , y = "Bill length (mm)", caption = "Source: Palmer Station LTER/palmerpenguins package", color = "Species") + scale_color_viridis_d()

```


# Mapping vs Setting

## Mapping flipper_length_mm to the alpha aesthetic.

```{r }
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm,
           size = body_mass_g,
           alpha = flipper_length_mm)) + 
  geom_point()

```

## Setting the alpha aesthetic to be 0.50.

```{r }
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm)) + 
  geom_point(size = 2, alpha = 0.5)

```

##  Is there any missing data? What is the plot doing with the missing values? Hint: consider using the skim() function from the skimr package to assess missingness.


```{r }
skim(penguins)
```

## Answer: Yes, there are missing data in Sex, bill_length_mm, bill_depth_mm, flipper_length_mm and body_mass_g. The ggplot() function, by default, will ignore missing values in the variables used for mapping aesthetics (x, y, and color in this case) and plot the available data points.

## To handle missing values explicitly, we would need to pre-process the dataset by removing or imputing missing values before creating the plot.

```{r }
# Remove rows with missing values
penguins_clean <- na.omit(penguins) #Now we have 333 obs instead of 344 

# Create the plot
ggplot(data = penguins_clean, mapping = aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
  geom_point() +
  labs(title = "Penguin bill length by depth",
       subtitle = "Dimensions for Adelie, Chinstrap, Gentoo penguins",
       x = "Bill depth (mm)",
       y = "Bill length (mm)",
       caption = "Source: Palmer Station LTER/palmerpenguins package",
       color = "Species")
```


