---
title: "Homework 6"
author: "Aditya Dube"
date: "`r format(Sys.time(), '%A, %B %e, %Y')`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Packages 

```{r }
library(stringr)
library(tidyverse)
library(flextable)
library(skimr)
library(ggplot2)
library(dplyr)
library(tidytext)
library(wordcloud2)
library(purrr)
library(janitor)
```


## Q1 Print and display the first 13 rows of standardDeck using the flextable() function from the flextable package.

```{r }
# Creating standard 52 card deck
cards <- c(as.character(2:10), "Jack", "Queen", "King", "Ace")

standardDeck <- expand_grid(suite = c("hearts", "diamonds", "clubs", "spades"),
                        value = cards) %>% 
  dplyr::mutate(value = as.ordered(fct_relevel(value, cards)))

# Displaying the first 13 rows using flextable
flextable(standardDeck[1:13,])
```

# Q2. Next, create a function called shuffle_cards() that randomly shuffles the rows of a tibble object representing a deck of cards. The shuffle_cards() function should have one argument, deck, a tibble with columns suite and value where each row corresponds to a card in the deck, and shuffle_cards() should return a tibble of the same structure, but with its rows randomly arranged / shuffled.

```{r }
shuffle_cards <- function(deck) {
  shuffled_deck <- deck %>% sample_n(nrow(deck))
  return(shuffled_deck)
}

shuffle_cards(standardDeck)
```
# Q3. We will now create a new function called deal_cards() that deals / gives nCards number of cards to a single player. The deal_cards() function should have 2 total arguments: deck, a tibble with columns suite and value where each row corresponds to a card in the card deck, and nCards, the total number of cards to be dealt / given to the player which should have a default value of 5.The deal_cards() function should return a tibble with columns suite and value and nCards number of rows.

```{r }
deal_cards <- function(deck, nCards = 5) {
  dealt_cards <- deck %>% sample_n(nCards)
  return(dealt_cards)
}

deal_cards(standardDeck, 7)

```

# Q4. Below are functions for detecting if a given hand has a flush, straight, straight flush, or royal flush. Copy and paste these functions for your own use. The hand argument should be a tibble with columns suite and value where each row corresponds to a card in a player’s hand. Create a function called detect_full_house() that has one argument, hand, and returns TRUE if hand has a full house and returns FALSE otherwise.


```{r }
# Create function to detect if a hand has a flush
detect_flush <- function(hand) {
  if(length(unique(hand$suite)) == 1) {
  return(TRUE)
  } else {
    return(FALSE)
  }
}

# Create function to detect if a hand has a straight
detect_straight <- function(hand) {
  
  # Unique cards
  ucards <- unique(hand$value)
  
  if(length(ucards) < 5) {
    return(FALSE)
  }
  
  # Sort the values of the hand in ascending order
  sortedHand <- sort(ucards)
  
  # Check if Ace is in hand, and if so, check for 'high' or 'low' straight
  if("Ace" == sortedHand[5]) {
    straight <- (sortedHand[1] == "10" & sortedHand[2] == "Jack" &
                sortedHand[3] == "Queen" & sortedHand[4] == "King") |
                (sortedHand[1] == "2" & sortedHand[2] == "3" &
                sortedHand[3] == "4" & sortedHand[4] == "5")
      
  } else {
  # Check if there are any gaps larger than 1 in the sequence of differences 
  straight <- !any(diff(as.integer(sortedHand)) > 1)
  }
  
  return(straight)
}

# Create function to detect if a hand has a straight-flush
detect_straight_flush <- function(hand) {
  if(detect_flush(hand) & detect_straight(hand)) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}

# Create function to detect if a hand has a royal-flush
detect_royal_flush <- function(hand) {
  if(detect_straight_flush(hand) & (min(hand$value) == "10")) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}

detect_full_house <- function(hand) {
  value_counts <- table(hand$value)
  
  if (length(value_counts) == 2) {
    if (any(value_counts == 2) && any(value_counts == 3)) {
      return(TRUE)
    }
  }
  
  return(FALSE)
}



```



# Q5. Create a function called detect_two_pair() that has one argument, hand, and returns TRUE if hand has a two pair and returns FALSE otherwise.


```{r }
detect_two_pair <- function(hand) {
  value_counts <- table(hand$value)
  pair_count <- sum(value_counts == 2)
  
  if (pair_count >= 2) {
    return(TRUE)
  }
  
  return(FALSE)
}

```



# Q6. Create a function called detect_n_of_kind() that has two arguments, hand, and nkind, where nkind = 2 for detecting a hand with at least one pair, nkind = 3 for detecting a hand with three of a kind, and nkind = 4 for detecting a hand with four of a kind. The detect_n_of_kind() function should return TRUE if hand has the specified n of a kind and return FALSE otherwise. Note that detect_n_of_kind(hand, nkind = 2) should return TRUE if hand has a single pair, a two pair, a three of a kind, or a four of a kind.

```{r }
detect_n_of_kind <- function(hand, nkind) {
  value_counts <- table(hand$value)
  kind_count <- sum(value_counts == nkind)
  
  if (kind_count >= 1) {
    return(TRUE)
  }
  
  return(FALSE)
}

```



# Q7.The function calculate_hand() below uses all of the functions we have created to determine the best hand given a set of five cards for a player. Use this function to complete the rest of the assignment.


```{r }
# Creating function to determine hand
calculate_hand <- function(hand) {
  
  straight <- detect_straight(hand)
  flush <- detect_flush(hand)
  pair <- detect_n_of_kind(hand, nkind = 2)
  
  # Royal flush
  if(straight & flush) {
    if(detect_royal_flush(hand)) {
      return("Royal flush")
    } else {
      return("Straight flush")
    }
  } else if(flush) {
    return("Flush")
  } else if(straight) {
    return("Straight")
  } else if(!pair) {
    return("High card")
  } else if(pair & detect_n_of_kind(hand, nkind = 4)) {
    return("Four of a kind")
  } else if(detect_full_house(hand)) {
    return("Full house")
  } else if(detect_n_of_kind(hand, nkind = 3)) {
        return("Three of a kind")
  } else if(detect_two_pair(hand)) {
        return("Two pair")
  } else {
        return("One pair")
  }
}

```

### Test Hands for Q7.

```{r }
testHands <- list()

# Should be a royal flush
testHands$`Royal flush` <- tibble(suite = rep("hearts", times = 5),
                 value = c("10", "Jack", "Queen",
                           "King", "Ace"))

# Should be a straight flush
testHands$`Straight flush` <- tibble(suite = rep("hearts", times = 5),
                 value = c("Ace", 2:5))

# Should be a four of a kind
testHands$`Four of a kind` <- tibble(suite = c(rep("hearts", 3), 
                                 rep("spades", 2)),
                 value = c(rep("Ace", 4), 7))

# Should be a straight
testHands$Straight <- tibble(suite = c(rep("hearts", 3), 
                          rep("spades", 2)),
                 value = c("Ace", 2:5))

# Should be a full house
testHands$`Full house` <- tibble(suite = c(rep("hearts", 3), 
                          rep("spades", 2)),
                 value = c(rep("Ace", 3), rep("King", 2)))

# Should be a flush
testHands$Flush <- tibble(suite = rep("hearts", times = 5),
                 value = c("Ace", 3:6))

# Should be a three of a kind
testHands$`Three of a kind` <- tibble(suite = c(rep("hearts", 3), 
                          rep("spades", 2)),
                 value = c(rep("Ace", 3), 2, 7))

# Should be a two pair
testHands$`Two pair` <- tibble(suite = c(rep("diamonds", 3), 
                          rep("clubs", 2)),
                 value = c(rep("Ace", 2), rep("Jack", 2), 7))

# Should be a one pair
testHands$`One pair` <- tibble(suite = c(rep("diamonds", 3), 
                          rep("clubs", 2)),
                 value = c(rep("Ace", 2), 2, 3, 7))

# Should be a high card
testHands$`High card` <- tibble(suite = c(rep("diamonds", 3), 
                          rep("clubs", 2)),
                 value = c("Ace", "Jack", 2, 3, 7))

# Testing hands
map_chr(.x = testHands, .f = function(hand) {
  calculate_hand(hand %>% 
  dplyr::mutate(value = as.ordered(fct_relevel(value, cards))))
  })

```


# Q8. Create a function called poker_sim() that has two arguments, seed, which specifies a seed value for reproducible random number generation, and deck, which we will supply with standardDeck when calling the function. The poker_sim() function should first set a seed value, shuffle the deck using the shuffle_cards() function, select 5 cards using the deal_cards() function, and then return the result of calculate_hand() on this dealt hand.

```{r }
poker_sim <- function(seed, deck) {
  set.seed(seed)
  shuffled_deck <- shuffle_cards(deck)
  dealt_cards <- deal_cards(shuffled_deck, nCards = 5)
  result <- calculate_hand(dealt_cards)
  return(result)
}

```


# Q9. Implement 10,000 poker hands using the poker_sim() function and the map_chr() function from the purrr package, storing the result of the simulations in a character vector called simRes.

```{r }
seed <- 123
num_simulations <- 10000

simRes <- map_chr(1:num_simulations, ~poker_sim(seed + .x, standardDeck))

```


# Q10. Summarize the simulations using the tabyl() function from the janitor package and the flextable function from the flextable package to produce a table structured as below. Note that your numbers will likely be slightly different.

```{r }
summary_table <- simRes %>%
  tabyl() %>%
  flextable()

summary_table
```



 
